Mobile Assets in Semantic Digital Twins

### 2023-05-19 paper/main.tex
Fix text, add backup of paper

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{comment}
\usepackage{duomasterforside}
\usepackage{listings}
\usepackage{float}
\usepackage{fancyvrb}
\usepackage{xcolor}
\usepackage{fvextra}
\usepackage{soul}
\usepackage{minted}
\graphicspath{ {./graphics/} }
\usepackage[backend=biber,natbib,style=numeric]{biblatex}
\usepackage{hyperref}
\addbibresource{references.bib}

@@

\section*{Abstract}

@@

\section*{Acknowledgements}
\newpage

@@

\listoftables
\newpage

@@

@@

It is stated by \citeauthor{godager_concept_2021} that although BIM could serve as a basis for DTs by providing lots of relevant information, DTs give even more context about the built environment, especially by monitoring assets from near real-time data. A lot of work goes into modeling physical assets in general \cite{waszak_let_2022}, and using BIM is also very costly due to frequent updates by maintainers \cite{hamledari_ifc-based_2018}.

\citeauthor{waszak_let_2022} proposes the use of knowledge graphs (i.e. using graphs to represent data from diverse sources \cite{hogan_introduction_2022, ryen_building_2022}) in the interconnection of various elements in DTs. Instead of storing all related data in one place, they propose storing static and dynamic data sources in different places and linking them.

Handling mobile assets in a DT based on a dynamic asset model (i.e. a file that contains an organized description of an asset's composition and properties \cite{kamburjan_twinning-by-construction_2022} is unexplored territory.

\subsection{Motivation}
Given the gap that exists in connecting tracking systems, asset models, and digital twins, there is still room for improvement. The problem is that there exist no systems that have a digital twin (DT) based on a dynamic asset model which can track an unlimited number of heterogeneous smartphones. In an attempt to formalize heterogeneous data and make it understood by computers, semantic web technologies should be used. More specifically, we aim to semantically enrich a model of a building to increase interoperability, as well as keep it inexpensive and easy to maintain.

We envision a \emph{semantic} digital twin (i.e. a digital twin that processes heterogeneous data from different sources in the domain into generally understandable information \cite{birgit_boss_digital_nodate}) based on a dynamic asset model in which static data (existing building infrastructure) and dynamic data (smartphone's physical location) is separated. The DT will be able to track an unlimited number of mobile assets simultaneously. The DT will be formalized using semantic web technologies. In the knowledge graph, static data will be noted to be different than dynamic data as it is generated from a combination of the program state and the asset model.

The asset model will be automatically reloaded by the system itself, or manually updated by an operator either offline (online version updated from local changes), or directly online in an ontology editor. This way the mobile assets are not only tracked, but the physical system containing them is twinned in near-real time as well. The DT will get positional sensor data from the PT, and then send informed decisions back to the PT.

\subsection{Problem statement}
From the context of digital twins and an increasing number of smartphone users, as well as the motivation for improvement, we present the following hypothesis (H):

@@

In addition to this, we present the following three research questions which will guide this thesis:
\begin{itemize}
    \item[\textbf{RQ1:}]
    Can we create a dynamic asset model of a simple building that is also extensible?
    \item [\textbf{RQ2:}] 
    Can we enable bidirectional data flow between the PT and DT, such that the digital twin gets updated sensor data from mobile devices and sends informed decisions back?
    \item [\textbf{RQ3:}]
    Can we clearly separate dynamic data (smartphone's position) from static data (building infrastructure) in the asset model and in the DT?
\end{itemize}

Having the aforementioned in mind, a possible overview of this thesis' components is shown below in Figure \ref{fig:initial_components}.

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.14]{graphics/initial_thesis_overview.png}
    \caption{Possible overview of architectural components in this thesis.}
    \label{fig:initial_components}
\end{figure}

\subsection{Thesis scope and outline}\label{subsec:Scope}
The scope of this thesis is to handle an unlimited number of mobile assets in a semantic digital twin based on a dynamic asset model. Figure \ref{fig:initial_components} shows the possible overview of architectural components and the bidirectional data flow between the PT and DT. The digital twin should be developed using the Semantic Micro Object Language (SMOL) described in \ref{subsec:SMOL}(SMOL) and its publicly available interpreter\footnote{\url{https://github.com/smolang/SemanticObjects}}.

@@

The areas are predefined squares with latitude and longitude ranges, where we assume the coordinates align. Square areas are easier to compute and scale into larger areas, and we are not concerned with other shapes such as polygons. The analysis of what certain area(s) are (e.g. what does a critical area mean) is also outside the scope of this thesis.

@@

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.3]{graphics/simple_building.png}
    \caption{A simple building with the structural components: two rooms (R1 and R2) connected by an inner wall (W) with a door (D) in it.}
    \label{fig:simple_building}
\end{figure}

The remainder of this thesis is organized as follows:
\begin{itemize}
    \item \hyperref[sec:Background]{\textbf{Background}} describes the theoretical background of this thesis.
    \item \hyperref[sec:Analysis]{\textbf{Problem analysis and design}} investigates the problem statement and research questions. We also look at what a dynamic asset model means, as well as the separation of static and dynamic data.
    \item \hyperref[sec:Implementation]{\textbf{Implementation}} dives into the implementation of the proposed solution and how the different architectural components shown in Figure \ref{fig:components} interact. More specifically, how do we "close the loop" by sending informed decisions back to the PT. 
    \item \hyperref[sec:Evaluation]{\textbf{Evaluation}} evaluates the proposed solution according to the structural requirements that were set.
    \item \hyperref[sec:Discussion]{\textbf{Discussion}} discusses the discoveries of this thesis and compares them to other existing results and theories. This section also explores the originality of the discoveries.
    \item \hyperref[sec:Conclusion]{\textbf{Conclusion and further work}} concludes this thesis. Mainly we look at the improvements that can be made to the proposed solution in the future.  
\end{itemize}

\newpage
\section{Background}\label{sec:Background}
This section describes the background information of this thesis. Starting off we give an in-depth description of digital twins (DTs). Then we give a description of the vision of a Semantic Web, as well as the technologies that help us build it. Later on we describe semantic reasoning and tools for it.  Then we look at frameworks for building cross-platform applications from a single codebase. Near the end of the section, we describe real-time and time-series databases.

\subsection{Digital Twin}\label{subsec:DigitalTwins}

The \emph{idea} of a digital twin can be traced back to the 1960s in NASA's space exploration days, where multiple simulations were employed to evaluate the issues with the space mission Apollo 13 \cite{noauthor_digital_nodate, fuller_digital_2020}. The \emph{term} digital twin was first introduced to the public by Michael Grieves in a presentation given in 2002 which was also later documented in \cite{grieves_michael_digital_2014} in 2014. This set the foundation for developing digital twins \cite{grieves_michael_digital_2014, fuller_digital_2020}.

As of today, there is no agreed-upon definition of a DT. However, it can be referred to as a digital representation of some physical system. There is no restriction on the size or scale of the physical system, which means the entity could be a single asset or a system of systems for that matter \cite{li_digital_2022, waszak_let_2022}. It exists in real life and is commonly referred to as the physical twin (PT), and is observed by the DT in near real-time. There should be a bidirectional flow of data between the DT and the PT throughout the life cycle of the PT. In other words, the DT gets updated sensor data from the PT, and sends informed decisions back to it \cite{madni_leveraging_2019, waszak_let_2022, kamburjan_digital_2022}.

@@

DTs have had many applications in various disciplines over the years. Although the idea of a digital twin was adopted early in the aerospace industry by NASA \cite{li_digital_2022}, later years have seen widespread use elsewhere \cite{fuller_digital_2020, waszak_let_2022, macchi_exploring_2018}, such as:

\begin{itemize}
    \item Smart Cities
    \item Healthcare
    \item Manufacturing
    \item Asset Management
\end{itemize}

There are many challenges with DTs. Some of these concern data, privacy, and costs of modeling physical assets \cite{fuller_digital_2020, waszak_let_2022}. General Data Protection Regulation (GDPR)\footnote{\url{https://gdpr-info.eu/}} works to ensure the privacy and security of personal data. As an example, no personal data should be collected without the permission of the \emph{data subject} (i.e. user), stated in GDPR Article 13 and 14 \cite{noauthor_guide_nodate}.

It is important to understand what a DT is and what it is not. We should try to "close the loop" by sending informed decisions back to the physical counterpart so that it reflects the newest state of the DT. A DS could be used alongside the DT with the only goal of being a timestamped snapshot of some specific physical asset \cite{bergs_concept_2021}. When implementing one should also think about the challenges of digital twins, such as privacy concerns and costs. We should not get too caught up with modeling an exact 3D replica of an asset. On the other hand, we \emph{should} use existing semantic technologies such as knowledge graphs for formal knowledge representation from a variety of data sources.

\subsection{Semantic Web}
Semantic Web was \emph{envisioned} by Sir Tim Berners-Lee as an extension of the World Wide Web (WWW) in which data was linked. Linked data is about making links so people and machines can explore the web of data, and according to Berners-Lee, should follow the four rules \cite{tim_berners-lee_linked_nodate} below:

\begin{enumerate}
    \item Uniform Resource Identifiers (URIs) should be used as names for things.
    \item HTTP URIs should be used so that people can look up the names.
    \item Useful information should be provided when people look up URIs, by using RDF or SPARQL ((SPARQL Query Language for RDF).
    \item Links to other URIs should also be provided so that people can discover more things.
\end{enumerate}

@@

\begin{itemize}
    \item \textbf{RDF} is a data model for describing (meta)data and its interchange on the web. Its linking structure forms a directed, labeled graph (RDF graph) \cite{sanga_spectral_2020, noauthor_sparql_nodate}. An RDF graph is a set of RDF triples. Triples consist of a subject, a predicate, and an object, as shown in Figure \ref{fig:rdf_triple_simple}. They can be thought of as facts. By using them, structured metadata can be exposed and shared between different applications \cite{noauthor_resource_nodate}. Triples are typically stored in triple stores (databases built for purposely storing triples) and can be retrieved with semantic queries \cite{noauthor_triple_nodate}.

    \begin{figure}[H]
        \centering
        \includegraphics[scale=0.16]{graphics/rdf_triple_simple.png}
        \caption{Example of an RDF triple consisting of a subject, a predicate, and an object.}
        \label{fig:rdf_triple_simple}
    \end{figure}

@@

    \begin{figure}[H]
        \centering
        \includegraphics[scale=0.18]{graphics/rdf_graph_example.png}
        \caption{RDF graph consisting of two triples in its set.}
        \label{fig:rdf_graph_example}
    \end{figure}
    \item \textbf{SPARQL}  is a query language for RDF and is defined by W3C. It expresses queries for varying sources of data, and the results from querying are sets of results or RDF graphs \cite{noauthor_sparql_nodate}.

    Below is an example of querying the title (\textbf{Building}) in the RDF graph shown in Figure \ref{fig:rdf_graph_example}.
\end{itemize}
\textbf{Data:}
\begin{Verbatim}[breaklines=true,breaksymbol=]
<http://www.semanticweb.org/oscarlr/ontologies/2023/2/building> 
<dc:title> 
”Building”
\end{Verbatim}
\textbf{Query:}
\begin{Verbatim}[breaklines=true,breaksymbol=]
SELECT ?title 
WHERE
{
  <http://www.semanticweb.org/oscarlr/ontologies/2023/2/building> 
  <dc:title> 
  ?title . 
}
\end{Verbatim}
\textbf{Result:}
\begin{table}[H]
    \begin{tabular}{|c|}
        \hline
        \textbf{title} \\
        \hline
        "Building" \\
        \hline
    \end{tabular}
    \caption{Query Result}
    \label{tab:my_label}
\end{table}

@@

\subsection{Asset Model}\label{subsec:AssetModel}
Although BIM (Building Information Modeling) can be used with RDF and JSON to create static building infrastructure as mentioned in \ref{sec:Introduction}(Introduction), some meaning is lost in translation due to a lack of semantic enrichment. BIM can be used in DTs for more context about the built environment which includes asset behavior \cite{godager_concept_2021}. However, a digital twin should aim to have interoperability between heterogeneous data sources.

OWL documents (i.e. ontologies) can easily be created and maintained in ontology editors, such as Protégé\footnote{\url{https://protege.stanford.edu}}. Such a tool can be used to give semantics to e.g. a model of a building, as well as spending less time on creating and maintaining it compared to BIM.

Figure \ref{fig:building_ontology} shows an example of an ontology of the simple building shown in Figure \ref{fig:simple_building} in Turtle Syntax (i.e. data format for RDF data model \cite{noauthor_terse_nodate}). This ontology uses semantic technologies outlined in Section \ref{subsubsec:Technologies}(Technologies), such as RDF, which again is used by OWL to formalise knowledge in this specific building domain. For a more compact view of the ontology, URLs are removed and the coordinates of the room border corners are purposely left out, and some white space and comments are also removed. We also do not take directions into account meaning the room R1 will always be left of R2, and R2 always to the right of R1. Note that this is just an example of an ontology and that a real building domain would be many times more complex.

@@

Lastly, this ontology's data can also be reasoned over with technologies outlined in \ref{subsec:Reasoning}(Reasoning), which contains an example of reasoning over an \emph{inconsistent} ontology in Protégé (desktop version of editor), as well as a description on how to make the ontology \emph{consistent}.

\begin{figure}[H]
    \centering
    \caption{Ontology of a simple building with the object properties (hasDoor, hasWallLeft, hasWallRight), the classes (Room, Wall, Door), and the individuals (D, R1, R2, W).}
    \label{fig:building_ontology}
    \begin{Verbatim}[frame=single]
ex:hasDoor rdf:type owl:ObjectProperty .
ex:hasWallLeft rdf:type owl:ObjectProperty .
ex:hasWallRight rdf:type owl:ObjectProperty .

@@

The source code of SMOL is publicly available\footnote{\url{https://github.com/smolang/SemanticObjects}}.

\subsubsection{Core Language}\label{subsubsec:CoreLanguage}
The \emph{lexical structure} (i.e. some basic rules for defining how code is written in a given language) of the SMOL language has a grammar that is defined using a simple \emph{EBNF notation} (i.e. notation for formally specifying syntax) as defined by W3C \cite{noauthor_lexical_nodate, noauthor_ebnf_nodate}. What follows is a simple SMOL program that mainly creates a smartphone object and prints \verb|Careful!| if a criterion is met:

\begin{figure}[H]
    \centering
    \caption{A simple SMOL program (file: example.smol)}
    \label{fig:smol_program}
    \begin{Verbatim}[frame=single]
class Smartphone(Int id, String name, Boolean isBlackBerry) end

@@

\subsubsection{SMOL Interpreter}\label{subsubsec:SMOLInterpreter}

The interpreter reads and executes the SMOL program in a given file. The example program (in the file \textbf{example.smol}) shown in Figure \ref{fig:smol_program} was executed in the terminal. Note that the second and fourth line starts with \verb|MO>|. This simply means commands are executed inside the REPL.
\begin{Verbatim}[frame=single]
java -jar smol.jar
MO> reada example.smol
Careful!
MO>
\end{Verbatim}

\subsection{Reasoning}\label{subsec:Reasoning}
OWL was previously mentioned in \ref{subsubsec:Technologies}(Technologies) as a language for expressing rich knowledge of concepts and relations between entities. We should keep in mind, however, that this formalised knowledge should be reasoned over and put in context to draw conclusions based on criteria. There are many technologies that let us do this.
\subsubsection{HermiT}
HermiT is a reasoner for OWL 2 \cite{glimm_hermit_2014}, which is an upgraded version of OWL. \citeauthor{glimm_hermit_2014} describes that the reasoner support both object and data property classification, as well as SPARQL query answering. HermiT is built into the ontology editor Protégé and can be used to reason over an ontology directly in the editor.

Figure \ref{fig:hermit_in_protege} shows an ontology that is \emph{inconsistent} (an ontology that cannot have any models and entails everything \cite{horridge_explaining_2009, huang_reasoning_2004}). The ontology is inconsistent because the object properties \textbf{hasWallLeft} and \textbf{hasWallRight} are disjoint (having no elements in common), and the individual \textbf{R1} has both object properties. This makes no sense because the room can't have the wall on its left side and at the same time have it on its right side. This could be fixed by simply removing the object property \textbf{hasWallLeft} from the individual according to the context. The effect of an inconsistent ontology is that no meaningful conclusions can be drawn from it \cite{horridge_explaining_2009}.

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.32]{graphics/screenshot_ontology.png}
    \caption{Shows use of HermiT to reason over an ontology directly in Protégé. The information window \textbf{Help for inconsistent ontologies} tells us that the OWL reasoner (HermiT) will no longer be able to provide any useful information about the ontology.}
    \label{fig:hermit_in_protege}
\end{figure}

@@

\subsection{App Development}\label{subsec:AppDevelopment}
Compared to previously covered topics, such as digital twins (DTs) and the Semantic web, app development is not a well-researched topic, as it is mainly used in industry. We will therefore take an industrial approach in describing the topic of app development.

Smartphones and tablets have different operating systems for mobile, in which the two most notable are iOS (iPhone and iPad) and Android (smartphones and tablets with different brands, such as Samsung, Google, and OnePlus). Smartphones are heterogeneous by nature. They not only have different operating systems but also have various device sensors, depending on the brand and price of the device.

@@

We should consider that although databases can have many different uses, some are better for specific purposes than others. Time series databases can be integrated into digital twins (DTs) for getting updated sensor data from the physical twin (PTs). \ref{subsec:SMOL}(SMOL) describes that SMOL can be used as a framework to create DTs. Additionally, in SMOL we can query time series data from InfluxDB by using the \verb|access| statement \cite{noauthor_time_nodate}. An example program of querying latitude and longitude coordinates from the start of the time range, and then returning the list of coordinates follows:

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{comment}
\usepackage{duomasterforside}
\usepackage{listings}
\usepackage{float}
\usepackage{fancyvrb}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{fvextra}
\usepackage{soul}
\usepackage{minted}
\usepackage{amssymb}
\usepackage{pifont}
\usepackage{multirow}
\graphicspath{ {./graphics/} }
\usepackage[backend=biber,natbib,style=numeric]{biblatex}
\usepackage{hyperref}
\addbibresource{references.bib}

@@

\section*{Abstract}
This thesis proposes a solution to the challenge of handling mobile assets in a semantic digital twin based on a dynamic asset model in which static data (existing building infrastructure) is separated from dynamic data (smartphone's physical location). The digital twin (DT) is created in Semantic Micro Object Language (SMOL) for easy integration with semantic technologies, such as semantic knowledge graphs. In order to semantically enrich the DT further, a model of a building is created in Web Ontology Language (OWL). A hybrid smartphone application is implemented having the mobile assets in mind, which sends their updated status. The privacy of the data subjects is protected as much as possible. The DT is an up-to-date digital representation of the PT, and it reconfigures in near real-time if the PT or any of the mobile assets inside it changes. If a person is located inside a dangerous area, the app warns the appropriate user from the informed decision sent by the DT.

@@

\section*{Acknowledgements}
I would like to extend my greatest gratitude to those that helped me with this thesis. Thanks to Einar Broch Johnsen, Rudolf Schlatte, and Eduard Kamburjan for formidable guidance in writing this thesis. Einar, thanks for believing in me from the start. I have learned a lot from the discussions on separating the static analysis from the dynamic snapshots. Rudi, I have thoroughly enjoyed our many discussions on relevant subjects to this thesis, especially about the building domain and privacy. Eduard, thanks for giving me great feedback on my writing and everything SMOL related.

Thanks to my family, and especially my parents, for being there for me no matter what. I could never have done this without any of you.      
\newpage

@@

@@

It is stated by \citeauthor{godager_concept_2021} that although BIM could serve as a basis for DTs by providing lots of relevant information, DTs should give even more context about the built environment, especially by monitoring assets from near real-time data. A lot of work goes into modeling physical assets in general \cite{waszak_let_2022}, and using BIM is also very costly due to frequent updates by maintainers \cite{hamledari_ifc-based_2018}.

 Knowledge graphs (i.e. using graphs to represent data from diverse sources \cite{hogan_introduction_2022, ryen_building_2022}) are used in the interconnection of various elements in DTs. Furthermore, instead of storing all related data in one place, \citeauthor{waszak_let_2022} propose storing static and dynamic data sources in different places and linking them.

@@

Handling mobile assets in a DT based on an \emph{asset model} (i.e. a file that contains an organized description of an asset's composition and properties \cite{kamburjan_twinning-by-construction_2022}) that is \emph{dynamic} is unexplored territory.

\subsection{Motivation}\label{subsec:Motivation}
Given the gap that exists in connecting tracking systems, asset models, and digital twins, there is still room for improvement. The problem is that there exist no systems that have a digital twin (DT) based on a dynamic asset model which can track an unlimited number of heterogeneous smartphones. In an attempt to formalize heterogeneous data and make it understood by computers, semantic web technologies should be used. More specifically, we aim to semantically enrich a model of a building to increase interoperability, as well as keep it inexpensive and easy to maintain.

We envision an extension of a \emph{semantic} digital twin (i.e. a digital twin that processes heterogeneous data from different sources in the domain into generally understandable information \cite{birgit_boss_digital_nodate}) that handles mobile assets, and is based on a dynamic asset model in which static data (existing building infrastructure) and dynamic data (smartphone's physical location) is separated. The DT will be able to track an unlimited number of mobile assets simultaneously. The DT will be formalized using semantic web technologies. In the knowledge graph, static data will be noted to be different than dynamic data as it is generated from a combination of the program state and the asset model.

The asset model will be automatically reloaded by the system itself, or manually updated by an operator either offline (online version updated from local changes), or directly online in an ontology editor. This way the mobile assets are not only tracked, but the physical system containing them is twinned in near-real time as well. The DT will get sensor data from the PT, and then send informed decisions back to the PT.

\subsection{Problem statement}\label{subsec:ProblemStatement}
From the context of digital twins and an increasing number of smartphone users, as well as the motivation for improvement described in Section \ref{subsec:Motivation}, we present the following hypothesis (H):

@@

In addition to this, we present the following three research questions which will guide this thesis:
\begin{itemize}
    \item[\textbf{RQ1:}]
    How to create a dynamic asset model of a simple building that is also extensible?
    \item [\textbf{RQ2:}] 
    How to enable bidirectional data flow between the PT and DT, such that the digital twin gets updated sensor data from mobile devices and sends informed decisions back?
    \item [\textbf{RQ3:}]
    How to separate dynamic data (smartphone's physical location) from static data (building infrastructure) in the asset model?
\end{itemize}

Having the aforementioned in mind, an overview of the main architectural components in this thesis is shown below in Figure \ref{fig:initial_components}.

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.14]{graphics/initial_thesis_overview.png}
    \caption{Overview of \emph{main} architectural components in this thesis}
    \label{fig:initial_components}
\end{figure}

\subsection{Thesis scope and outline}\label{subsec:Scope}
The scope of this thesis is to handle an unlimited number of mobile assets in a semantic digital twin based on a dynamic asset model. Figure \ref{fig:initial_components} shows the overview of architectural components and the bidirectional data flow between the PT and DT. The digital twin should be developed using the Semantic Micro Object Language (SMOL) described in Section \ref{subsec:SMOL} and its publicly available interpreter\footnote{\url{https://github.com/smolang/SemanticObjects}}.

@@

The areas are predefined squares with latitude and longitude ranges, where we \emph{assume} the coordinates align. Square areas are easier to compute and scale into larger areas, and we are not concerned with other shapes such as polygons. The analysis of what certain area(s) are (e.g. what does a critical area mean) is also outside the scope of this thesis.

@@

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.3]{graphics/simple_building.png}
    \caption{A simple building with the structural components: two rooms (R1 and R2) connected by an inner wall (W) with a door (D) in it}
    \label{fig:simple_building}
\end{figure}

The remainder of this thesis is organized as follows:
\begin{itemize}
    \item Chapter \ref{sec:Background} describes the theoretical background of this thesis.
    \item Chapter \ref{sec:Analysis} investigates the problem statement and the research questions. We also look at what a dynamic asset model means, as well as the separation of static and dynamic data.  
    \item Chapter \ref{sec:Implementation} dives into the implementation of the proposed solution and how the different architectural components shown in Figure \ref{fig:components} interact. More specifically, how do we "close the loop" by sending informed decisions back to the PT. 
    \item Chapter \ref{sec:Evaluation} evaluates the proposed solution according to the structural requirements that were set, and then according to various test runs.
    \item Chapter \ref{sec:Discussion} discusses the limitations of the discoveries in this thesis, the alternatives, and compares them to other existing work.
    \item Chapter \ref{sec:Conclusion} concludes this thesis and presents the contributions. We look at improvements that can be made to the proposed solution, and further work that can be done using the thesis as a basis.  
\end{itemize}

\newpage
\section{Background}\label{sec:Background}
This chapter describes the background information of this thesis. Starting off we give a description of digital twins (DTs). Then we give a description of the Semantic Web, as well as the technologies that help us build it. Later on we describe semantic reasoning and tools for it. Then we look at frameworks for building cross-platform applications from a single codebase. Near the end of the section, we describe real-time and time-series databases.

\subsection{Digital Twin}\label{subsec:DigitalTwins}
\subsubsection{Historic background}
The \emph{idea} of a digital twin can be traced back to the 1960s in NASA's space exploration days, where multiple simulations were employed to evaluate the problems with the space mission Apollo 13 \cite{noauthor_digital_nodate, fuller_digital_2020}. The \emph{term} digital twin was first introduced to the public by Michael Grieves in a presentation given in 2002, and in 2014 conceptualized as a virtual representation of a factory replication \cite{grieves_michael_digital_2014}. This set the foundation for developing digital twins \cite{grieves_michael_digital_2014, fuller_digital_2020}. Since then, DTs have had many applications in various disciplines, and in recent years especially in smart cities, healthcare, manufacturing, and asset management \cite{fuller_digital_2020, waszak_let_2022, macchi_exploring_2018}

\subsubsection{For the scope of this thesis}
As of today, a DT is referred to as a digital representation of some physical system \cite{grieves_digital_2017}, that has no restriction on the size or scale of it, which means the entity could be a single asset or a system of systems for that matter \cite{li_digital_2022, waszak_let_2022}. It exists in real life and is commonly referred to as the physical twin (PT), and is observed by the DT in near real-time. There should be a bidirectional flow of data between the DT and the PT throughout the life cycle of the PT. In other words, the DT gets updated sensor data from the PT, and sends informed decisions back to it \cite{madni_leveraging_2019, waszak_let_2022, kamburjan_digital_2022}. From this, and for the scope of this thesis, a DT is a digital representation of a system that tracks many mobile assets. The DT sends informed decisions to the physical counterpart in which the appropriate mobile assets perform actions.

@@

To sum up, it is important to understand what a DT is and what it is not. We should try to "close the loop" by sending informed decisions back to the physical counterpart so that it reflects the newest state of the DT. A DS could be used alongside the DT with the only goal of being a timestamped snapshot of some specific physical asset \cite{bergs_concept_2021}. When implementing one should also think about the challenges of digital twins, such as privacy concerns and costs. We should not get too caught up with modeling an exact 3D replica of an asset. On the other hand, we \emph{should} use existing semantic technologies such as knowledge graphs for formal knowledge representation from a variety of data sources.

\subsection{Semantic Web}
Semantic Web was \emph{envisioned} as an extension of the World Wide Web (WWW) in which data was linked. Linked data is about making links so people and machines can explore the web of data, and according to Tim Berners-Lee, should follow the four rules \cite{tim_berners-lee_linked_nodate} below:

\begin{enumerate}
    \item Uniform Resource Identifiers (URIs) should be used as names for things.
    \item HTTP URIs should be used so that people can look up the names.
    \item Useful information should be provided when people look up URIs, by using RDF or SPARQL.
    \item Links to other URIs should also be provided so that people can discover more things.
\end{enumerate}

@@

\begin{itemize}
    \item \textbf{RDF} is a data model for describing (meta)data and its interchange on the web. Its linking structure forms a directed, labeled graph (RDF graph) \cite{sanga_spectral_2020, noauthor_sparql_nodate}. An RDF graph is a set of RDF triples. Triples consist of a subject, a predicate, and an object, as shown in Figure \ref{fig:rdf_triple_simple}. They can be thought of as facts. By using them, structured metadata can be exposed and shared between different applications \cite{noauthor_resource_nodate}. Triples are typically stored in triple stores (databases purposely built for storing triples) and can be retrieved with semantic queries \cite{noauthor_triple_nodate}.

    \begin{figure}[H]
        \centering
        \includegraphics[scale=0.16]{graphics/rdf_triple_simple.png}
        \caption{Example of an RDF triple consisting of a subject, a predicate, and an object}
        \label{fig:rdf_triple_simple}
    \end{figure}

@@

    \begin{figure}[H]
        \centering
        \includegraphics[scale=0.18]{graphics/rdf_graph_example.png}
        \caption{RDF graph consisting of two triples in its set}
        \label{fig:rdf_graph_example}
    \end{figure}
    \item \textbf{SPARQL}  is a query language for RDF and is defined by W3C. It expresses queries for varying sources of data, and the results from querying are sets of results or RDF graphs \cite{noauthor_sparql_nodate}.

    Below is an example of querying the title (\textbf{Building}) in the RDF graph shown in Figure \ref{fig:rdf_graph_example}.
\end{itemize}
\textbf{Data:}
\begin{Verbatim}[breaklines=true,breaksymbol=]
<http://www.semanticweb.org/oscarlr/ontologies/2023/2/building> 
<dc:title> 
”Building”
\end{Verbatim}
\textbf{Query:}
\begin{Verbatim}[breaklines=true,breaksymbol=]
SELECT ?title 
WHERE
{
  <http://www.semanticweb.org/oscarlr/ontologies/2023/2/building> 
  <dc:title> 
  ?title . 
}
\end{Verbatim}
\textbf{Result:}
\begin{table}[H]
    \begin{tabular}{|c|}
        \hline
        \textbf{title} \\
        \hline
        "Building" \\
        \hline
    \end{tabular}
    \caption{Query result}
    \label{tab:my_label}
\end{table}

@@

\subsection{Asset Model}\label{subsec:AssetModel}
Although BIM (Building Information Modeling) can be used with RDF and JSON to create static building infrastructure as mentioned in Section \ref{sec:Introduction}, some meaning is lost in translation due to a lack of semantic enrichment. BIM can be used in DTs for more context about the built environment which includes asset behavior \cite{godager_concept_2021}. However, a digital twin should aim to have interoperability between heterogeneous data sources.

OWL documents can easily be created (from export), loaded into, and maintained in ontology editors, such as Protégé\footnote{\url{https://protege.stanford.edu}}. Such a tool can be used to give semantics to e.g. a model of a building, as well as spending less time on creating and maintaining it compared to BIM.

Figure \ref{fig:building_ontology} shows an example of an ontology of the simple building shown in Figure \ref{fig:simple_building} in Turtle Syntax (i.e. data format for RDF data model \cite{noauthor_terse_nodate}). This ontology uses semantic technologies outlined in Section \ref{subsubsec:Technologies}, such as RDF, which again is used by OWL to formalise knowledge in this specific building domain. For a more compact view of the ontology, URLs are removed and the coordinates of the room border corners are purposely left out, and some white space and comments are also removed. We also do not take directions into account meaning the room R1 will always be left of R2, and R2 always to the right of R1. Note that this is just an example of an ontology and that a real building domain would be many times more complex.

@@

Lastly, this ontology's data can also be reasoned over with technologies outlined in Section \ref{subsec:Reasoning}, which contains an example of reasoning over an \emph{inconsistent} ontology in Protégé (desktop version of editor), as well as a description on how to make the ontology \emph{consistent}.

\begin{figure}[H]
    \centering
    \caption{Ontology of a simple building with the object properties (hasDoor, hasWallLeft, hasWallRight), the classes (Room, Wall, Door), and the individuals (D, R1, R2, W)}
    \label{fig:building_ontology}
    \begin{Verbatim}[frame=single]
ex:hasDoor rdf:type owl:ObjectProperty .
ex:hasWallLeft rdf:type owl:ObjectProperty .
ex:hasWallRight rdf:type owl:ObjectProperty .

@@

The source code of SMOL is publicly available\footnote{\url{https://github.com/smolang/SemanticObjects}}. What follows is a simple SMOL program that creates a smartphone object and prints \verb|Careful!| if a criterion is met:

\begin{figure}[H]
    \centering
    \caption{A simple SMOL program}
    \label{fig:smol_program}
    \begin{Verbatim}[frame=single]
class Smartphone(Int id, String name, Boolean isBlackBerry) end

@@

\subsubsection{SMOL Interpreter}\label{subsubsec:SMOLInterpreter}
The interpreter reads and executes the SMOL program in a given file. The example program (in the file \textbf{example.smol}) shown in Figure \ref{fig:smol_program} was executed in the terminal. Note that the second and fourth line starts with \verb|MO>|. This simply means commands are executed inside the REPL.
\begin{Verbatim}[frame=single]
java -jar smol.jar
MO> reada example.smol
Careful!
MO>
\end{Verbatim}

\subsection{Reasoning}\label{subsec:Reasoning}
OWL was previously mentioned in Section \ref{subsubsec:Technologies} as a language for expressing rich knowledge of concepts and relations between entities. We should keep in mind, however, that this formalized knowledge should be reasoned over and put in context to draw conclusions based on criteria. There are many technologies that let us do this.
\subsubsection{HermiT}
HermiT is a reasoner for OWL 2 \cite{glimm_hermit_2014}, which is an upgraded version of OWL. \citeauthor{glimm_hermit_2014} describes that the reasoner support both object and data property classification, as well as SPARQL query answering. HermiT is built into the ontology editor Protégé and can be used to reason over an ontology directly in the editor. Other plugins for reasoning can be added to the ontology editor, but it is convenient that HermiT comes with the installation.

Figure \ref{fig:hermit_in_protege} shows an ontology that is \emph{inconsistent} (an ontology that cannot have any models and entails everything \cite{horridge_explaining_2009, huang_reasoning_2004}). The information window \textbf{Help for inconsistent ontologies} tells us that the OWL reasoner (HermiT) will no longer be able to provide any useful information about the ontology. The ontology is inconsistent because the object properties \textbf{hasWallLeft} and \textbf{hasWallRight} are disjoint (having no elements in common), and the individual \textbf{R1} has both object properties. This makes no sense because the room can't have the wall on its left side and at the same time have it on its right side. This could be fixed by simply removing the object property \textbf{hasWallLeft} from the individual according to the context. The effect of an inconsistent ontology is that no meaningful conclusions can be drawn from it \cite{horridge_explaining_2009}.

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.30]{graphics/screenshot_ontology.png}
    \caption{Shows use of HermiT to reason over an ontology in Protégé}
    \label{fig:hermit_in_protege}
\end{figure}

@@

\subsection{App Development}\label{subsec:AppDevelopment}
We will describe the topic of app development by looking at concerns when developing in industry. Smartphones and tablets have different operating systems for mobile, in which the two most notable are iOS (iPhone and iPad) and Android (smartphones and tablets with different brands, such as Samsung, Google, and OnePlus). Smartphones are heterogeneous by nature. They not only have different operating systems but also have various device sensors, depending on the brand and price of the device.

@@

We should consider that although databases can have many different uses, some are better for specific purposes than others. Time series databases can be integrated into digital twins (DTs) for getting updated sensor data from the physical twin (PTs). Section \ref{subsec:SMOL} describes that SMOL can be used as a framework to create DTs. Additionally, in SMOL we can query time series data from InfluxDB by using the \verb|access| statement \cite{noauthor_time_nodate}. An example program of querying latitude and longitude coordinates from the start of the time range, and then returning the list of coordinates follows:
